---
title: "Private Welfare in the Workplace and Support for Social Insurance"
subtitle: "the case of employee harship funds"
author: 
   - John S. Ahlquist^[Professor, School of Global Policy and Strategy, UC San Diego. jahlquist@ucsd.edu.  This project was pre-registered under osf.io/fhq93. This research received grant support from the Russell Sage Foundation, UCSD's Cowhey Center for Global Transformation, and the UCSD Yankelovitch Center.  Keng-Chi Chiang, Ofri Mantell, and Yuchen Wang provided research assistance.  Versions of this research were presented at the APSA, LERA, CAPE and EPSA meetings as well as seminars at Columbia, Duke, IE-Madrid, Universitat de Barcelona, Wisconsin, and Yale. Kate Bronfenbrenner, Daniel Galvin, Jake Grumbach, Margaret Levi, Elizabeth Lyons, and Eric Thai provided helpful comments.  Shane Xuan assisted with Facebook.]
   - Theodoros Ntounias^[PhD Candidate, department of political science, UC San Diego]
  
date: "`r paste(format(Sys.time(), '%d %B, %Y'), 'INCOMPLETE WORKING DRAFT', sep=' ')`"

output: 
  bookdown::pdf_document2:
    latex_engine: xelatex
    keep_tex: true
    toc: false
    #citation_package: natbib
    extra_dependencies: ["hyperref", "bbm", "subfig", "flafter"]
#biblio-style: "`r here::here('inputs', 'apsr.bst')`"
bibliography: 
   - "`r here::here('inputs', 'EHFbib3.bib')`"
   - "`r here::here('inputs', 'EHFbib4.bib')`"
link-citations: true
classoption: [11pt, oneside]
header-includes:
   - \usepackage{setspace}\onehalfspacing

---
\begin{abstract}
\begin{singlespace} 

There is a long-standing claim that public social insurance and welfare crowd out private savings and charity. There is also a newer set of arguments that improved access to private resources (home equity, credit) dampen voters' support for public insurance programs and redistributive taxation. We investigate this "substitution hypothesis" in the context of a novel set of workplace mutual aid arrangements--Employee Hardship Funds (EHFs). EHFs enable employers to distribute emergency cash to their workers using donations pooled from employees themselves. Hundreds of major US firms now operate these programs. We use three original surveys---including two matched to workers at specific employers with EHFs and including experiments.  We find that state-level generosity in the social safety net is unrelated to worker knowledge of and support for EHFs. Experimental manipulation of EHF salience among workers at two large retailers indicates that EHFs do not displace support for public unemployment insurance or government emergency aid, regardless of the generosity of the EHF program. [Something about political ideology]. EHFs, as a workplace mutual aid arrangement, and government social insurance programs appear unconnected in the minds of workers.\\


\end{singlespace}
\end{abstract}

```{r setup, include=FALSE}
#loading THD, WMT, and genpop data
temp<-here::here('2_code', '1_libraries_and_settings_global.R')
source(file = temp, local = knitr::knit_global())

temp<-here::here('2_code', 'THD', '2_data_format.R')
source(file = temp)

#WMT data with sample descriptives for appendix.
temp<-here::here('2_code','Walmart', '5_sample_desc.R')
source(file = temp)
rm(temp)

genpop <- read_csv(here("3_cleaned_data", "general_retail_clean.csv"))

genpop_w<- genpop  |>  
  as_survey_design(ids = 1, weights = acs_weight_trim)

```


# Introduction

# The "substitution hypothesis," two ways

[Lit review + theory]

# EHFs

[background and relevance]

# Study 1
## Research design
## Survey & Measurement
## Results
Genpop:
 EHF awareness/support/donation ~ state generosity (+ income + home ownership + past use of welfare benefits + partisanship/ideology/Trump + age + race + gender + ed + marital status + religiousity )

# Studies 2 & 3
## Case selection, justification, common survey and research design components
## Study 2
### Specifics of survey details
### Awareness and Donation willingness
EHF awareness/donation ~ state generosity (+ income + home ownership + past use of welfare benefits + partisanship/ideology/Trump + job tenure  + religiousity )

### Survey experiments
#### gov't aid/responsibility
 SI support ~ treatment (* awareness) (+ home ownership + past use of welfare benefits + partisanship/ideology/Trump + job tenure + age + race + gender + ed + marital status + religiousity )

#### COVID experiment (?)

## Study 3
### Specifics of survey details
### Awareness and Donation willingness
EHF awareness/donation ~ state generosity (+ income + home ownership + past use of welfare benefits + partisanship/ideology/Trump + job tenure + religiousity )

### Survey experiment
#### gov't aid/responsibility
 SI support ~ treatment (* awareness) (+ home ownership + past use of welfare benefits + partisanship/ideology/Trump + job tenure + age + race + gender + ed + marital status + religiousity )

#### donation action
 click donate ~ treatment (* awareness) (+ home ownership + past use of welfare benefits + partisanship/ideology/Trump + job tenure + age + race + gender + ed + marital status + religiousity )


# Discussion and Conclusion


# Appendix
## Deviation from PAP
## Survey details
## Alternative specifications & Study 1
## Alternative specifications & Study 2
## Manipulation check Study 3
## Alternative specifications & Study 3








Past generations of workers built organizations to help one another in hard times.  Some of these mutual aid societies eventually transformed into full-fledged trade unions.  Others morphed into mutual insurance companies that later informed government social insurance programs [@deswaanWorkersClientsMutualism1986].  Such organizations transformed workplace relations, socialized workers into political life, and produced labor-based political movements. 

Employers have tremendous leeway to introduce or modify benefit plans and workplace programs that may influence workers perceptions, attitudes, and behaviors including around actions like unionization. EHFs are one such tax-incentivized program that has received virtually no systematic study.  This oversight is all the more remarkable as EHFs clearly echo both an earlier era of welfare capitalism and past mutual aid initiatives that produced and modern labor unions.

The lack of any substitution effect between EHFs and support for public assistance suggests that EHFs may represent a reconfiguration of the moral economy of low-wage work, a recognition of the public system's many failings.  Rather than a substitute for social insurance and assistance, EHFs are low-cost signals of corporate values and selective benevolence.

<!---
######


In the contemporary United States, unions are vanishing and the welfare and social insurance systems are creaking. Temporary and other "non-standard" employment relationships are increasingly widespread [@abraham_measuring_2018].  Yet such work arrangements fail to provide access to traditional social insurance and labor market protections. Americans in the lower half of the education and earnings distributions have seen years of stagnating wages while also bearing more economic risk [@hacker_great_2019].  Precarious employment, "fissured" workplaces, and distributed supply chains make it more difficult for workers to develop the occupational identities and dense networks of coworkers that supported the organization-building of the past [@ahlquist_research_2018;@thelen_american_2019;@naidu_is_nodate;@weil_fissured_2014].  

Even as unions and social protections atrophy, employers across the U.S. are sponsoring a new form of private mutual aid: Employee Hardship Funds (EHFs). These tax-exempt programs distribute emergency cash grants to workers, funded in part through donations raised from the workers themselves. Although EHFs recall earlier eras of "welfare capitalism" [@brandesAmericanWelfareCapitalism1976;@jacobyModernManorsWelfare1998], these programs have gone largely unstudied. What do workers know about them? Can they influence job attachment and support for unionization?  What does this tell us about the "moral economy" between low-wage workers and employers in the US today?

EHFs present a tractable, unexploited context to study important questions in the social sciences and management.  EHFs present a well-defined intervention for studying  whether "private governance" initiatives actually benefit workers themselves [@anderson_private_2017; @ahlquist_firm_2020; @locke_promise_2013; @malesky_chains_2018; @distelhorst_does_2017; ].      

I argue that EHFs are a blend of employer-driven welfare capitalism and worker mutual aid.  Drawing on interviews with executives and EHF managers, I expect EHFs to be closely connected to the corporate culture in which they are embedded, producing mixed and contingent effects on worker attitudes. Using three original, pre-registered studies, I present the first _systematic_ description of worker awareness and experience with EHFs as well as experimental evidence on the effect of corporate messaging around EHFs. I show that EHFs can build attachment to employers and coworkers when they are generous, well-promoted, and align with other corporate signals and management practices, but they do not affect workers' subjective financial security.

In study 1, I use a survey of US retail workers to describe retail workers' awareness of EHF programs among major employers, their level of interest in such programs, and their preferred influence over how the fund is run.  I find that awareness is limited and retail workers are not well-informed about which firms offer such programs.  Nevertheless, when weighed against other job amenities, emergency cash support is second only to paid time off in inducing workers to consider a hypothetical job offer at another retailer.  Respondents strongly support the creation of an EHF at their current employer and preferred a program in which workers themselves have meaningful influence over fund management.

In studies 2 and 3,  I focus on workers at two large retailers with long-standing EHF programs: the Home Depot and Walmart.  Both firms have EHFs that are over two decades old and managed in-house.   But the firms and programs differ in important ways.  The Home Depot EHF, the Homer Fund, is considerably more generous (with a maximum grant of \$10,000 against Walmart's \$1,500) and more heavily promoted among workers and executives.  IRS filing data indicates that the Homer Fund is also considerably larger than Walmart's EHF, even though Walmart is over three times the size of the Home Depot in US employee headcount.  Both firms also maintain strikingly different corporate cultures, with the Home Depot known for.^[According to the MIT Sloan Culture 500 rankings, based on analysis of Glassdoor postings, Walmart's]  

Using social media to manage the difficult task of recruiting a sample of workers linked to specific firms, I fielded original, independent surveys of Home Depot and Walmart workers. I find considerable differences in program awareness across employers.  Among Home Depot workers, awareness is "broad-but-thin": over 60% of Home Depot respondents displayed some awareness of the Homer Fund, but there were inconsistencies across measurement items, suggesting recognition but not real understanding. At Walmart, only 38% were aware of the program.  At both employers, tenure with the firm is the primary consistent predictor of EHF awareness and engagement.

In both studies 2 and 3, I use survey experiments to identify the impact of messages that raise EHF awareness.  Specifically, I exploit employer-produced promotional video messages about the EHF and look at effects on workers' subjective financial security, job attachment, and support for unionization and government-provided social insurance. In the Walmart survey I look at additional outcomes, including willingness to click on a link to donate to the EHF and perceptions about co-workers' support for unionization. 

In Study 2, Home Depot workers who were already aware of the Homer Fund prior to the survey were more confident in their financial stability, more attached to their employer and coworkers, less supportive of unions, and less supportive of unemployment benefits, even after adjusting for job tenure.  The video treatment produced an average increase in job attachment of about 0.6 standard deviations on the relevant scale.  On average, the video treatment makes workers more _uncertain_ about unionization.  Further decomposing treatment effects, I find that the Homer Fund video treatment effects were largely concentrated among respondents _unaware_ of the EHF prior to my survey. Among this group, the video treatment significantly reduced support for unionization while increasing uncertainty (as opposed to outright opposition).

In Study 3, I find both similarities and differences with Study 2.  As in the Home Depot sample, Walmart workers who were already aware of their EHF prior to the survey were more confident in their financial stability, more attached to their employer and coworkers, and less supportive of unions, even after adjusting for job tenure.  For the experiment in Study 3, I construct a custom video vignette that reproduced the Homer Fund video for the Walmart context.  The video treatment again produced no detectable effect on perceived financial stability.  But, contrary to findings in the Home Depot sample, I find no evidence of a treatment effect on job attachment, including quit intentions.  The video treatment did produce a four percentage point increase in the likelihood a respondent takes an action to donate.  The video treatment reduces opposition to unionization and increases (by four percentage points) the percent of coworkers expected to vote for unionization.   In contrast to the Home Depot sample, the Walmart sample showed no evidence of heterogeneous treatment effects based on prior knowledge of the Walmart EHF, known as the Associates in Critical Need Trust Together Fund (ACNT). 
--->


From a policy perspective, EHFs appear to resonate with a felt need among low-wage retail workers in the US today.  But, as the Home Depot and Walmart surveys indicate, program details, context, and promotion matter for the effects they might have on the workforce.  These programs operate in the shadow of eroded public welfare institutions, yet they are funded in part by favorable tax treatment.  Are the benefits these programs provide worth the tax expenditures?  How do these programs change when workers themselves have a voice in how they are run? 

Workers value emergency aid arrangements, which coincides with the renewed emphasis on mutual aid--and other solidaristic selective incentives--as a basis for rebuilding worker solidarity [@hertel-fernandez_why_nodate; @jarley_unions_2005; @horowitz_mutualism_2021].  The development, management, and outreach around programs like EHFs provide a clear point of common interest between workers, unions, and management.  To date this possibility remains unexplored.


nor do they displace support for public benefits.


EHFs present a fascinating window into employer behavior and the state of the low wage labor market in the US and beyond. Future research can extend in a variety of productive directions.  


and more supportive of unemployment insurance

There is no detectable treatment effect on attitudes towards government programs in support of the unemployed or those facing hardship.

There is now a growing body of scholarship arguing that government welfare programs and private resources substitute for one another [@beito_mutual_2000; @yeo_working_1979; @wiedemann_how_2022].  Public programs often step in when private mutual aid or insurance pools collapse.  For example, the famous "Ghent system," in which unions administer government-mandated unemployment insurance, arose out of the repeated collapse of various privately-organized welfare and mutual aid pools [@socialsecurityadministrationSocialSecurityHistory2023; @western_1997].  @panjwaniCausesConsequencesMedical2023 show that health-related crowdfunding GoFundMe campaigns declined after expansions in medical insurance coverage due to the  Affordable Care Act.  @leeNewDigitalSafety2022 show that areas with poor private insurance coverage and fewer social associations see greater initiation rates for GoFundMe campaigns, but campaigns in these same areas are less likely to be successful.   Conversely, increases in private resources, private insurance, or employer-provided benefits can crowd out support for government programs [@ansell_political_2014; @busemeyer_welfare_2020; @rosner_struggle_2003; @hacker_insecure_2013; @zhu_policy_2015] .  Insofar as EHFs are a form of private emergency aid, increased program awareness could blunt support for social insurance or government aid programs in times of emergency (e.g., COVID-19).  This paper seeks to determine whether we can observe this effect among workers exposed to EHF information.

@hadziabdicPrivateSpannerPublic2022

EHF may be viewed as "private insurance", which could dampen support for social insurance. Prompting workers with information about their employer's EHF will reduce workers' willingness to vote for unionization and reduce stated support for government aid programs for unemployment or emergencies.


@fishbackSocialInsurancePublic2020

 attitudes, or beliefs about the welfare state?
 
 Some EHFs appear to be the result of spontaneous initiatives from executives, often in response to a major natural disaster affecting a region closely connected to the firm and its employees.^[Disaster aid and social insurance are closely linked in the US [@landis_fate_1999].]  Based on interviews, this often stemmed from frustration with the slow and uneven government aid. 

Even before COVID-19, there was plenty of evidence that the existing social insurance and welfare systems were stigmatizing, encumbered with numerous "administrative burdens" [@herd_administrative_2019], and straining to deliver to those in need.  Several informants suggested that EHFs may present a private, partial complement to a weakening or slow public system.  By locating mutual aid programs in the firm, program participation might be higher and the EHF can benefit from corporate resources.  Applying for an EHF grant may be easier or less stigmatizing than applying for government assistance and the turn around time for benefits could be much faster in times of acute need.^[The Home Depot claims that grants are processed within 7 business days of application receipt, far faster than welfare or unemployment benefits, which also require waiting periods in many states.]



 


### Attitudes toward government social insurance

In examining social insurance, I take a question battery from the General Social Survey that asks the extent to which the respondent thinks the government has a responsibility to provide for different constituencies.  Here I examine responses for "the unemployed" (4-category response ranging from "none at all" to "a lot").

Compared to unionization, workers in this sample are far more supportive of unemployment insurance, with a solid majority viewing the government as at least somewhat responsible for the welfare of the unemployed.  We see that the video treatment substantially reduces the proportion of responses in the "none" category in favor of the two positive categories.  The EHF video treatment made respondents more rather than less supportive of government-provided social insurance, contrary to expectations.  The text treatment, on the other hand, does not appear to have any detectable treatment effect on average.

```{r tab-ui,eval=TRUE, echo = FALSE, tab.cap="Support for UI regression models" }
temp<-here::here('2_code', 'THD', '9_ui_models.R')
source(temp)
rm(temp)
ui_tab

```

Table \@ref(tab:tab-ui) displays regression estimates for UI support.  Results here confirm a positive average treatment effect for the video treatment of about `r round(coef(ui_lm_uw)[3]/sd(THD_comp_uw$gov_ui_num,na.rm=T),2)` standard deviations.  In contrast to all of the outcomes studied above, we see no large or statistically discernible difference between the pre-exposed and unaware in average support for UI.  We also see no treatment effects among the unaware, although there are treatment effects among the _pre-exposed_, especially for the video treatment.

To interpret this more easily, Figure \@ref(fig:fig-uieff) displays treatment effects by respondent pre-exposure to the EHF.  In the case of UI, we see that treatment effects are entirely concentrated among those _already_ aware of the Home Depot's EHF.  Contrary to expectations, both the text and video treatments _increase_ support for UI, although only the latter is a significant effect by conventional standards.  This finding deserves more exploration, but it consistent with author interviews with EHF professionals in which they indicate that EHFs are meant to supplement or provide a bridge to public benefits, rather than supplant them.   

```{r fig-uieff, eval=TRUE, echo=FALSE, fig.cap="Interpreting treatment effects from UI models"}
ui_eff_plot

```

Is unemployment insurance the appropriate policy to link with the EHF?  After all, the EHF is designed to provide benefits while having a job with this employer whereas UI is designed to help when a job is lost.^[The pre-exposed may better appreciate this distinction, which may help explain the unexpected findings.]  There are two pieces of evidence that UI is a reasonable policy domain to link with EHF benefits.  First, although EHF benefits are not available to someone who loses her job with an employer, benefits are often available when a worker's immediate family member loses a job.  Second, the UI question was asked as part of a battery of questions asking about UI was well as two other policy areas: pensions for the elderly and the provision of childcare for working parents.  These two questions can constitute "placebos" since EHFs have no connection with either policy area.  Parameter estimates from regressing support for these alternative policy areas on the EHF treatments and EHF awareness appear in Appendix Table \@ref(tab:tab-placebo).  I find no evidence for any relationship between the EHF treatments and support for government involvement in either old age pensions or childcare, regardless of whether the respondent was aware of the EHF going into the survey.

Finally, we look at social insurance policies.  In the Walmart survey we again asked respondents the level of responsibility the government has for the elderly and the unemployed.  To further directly examine the substitution hypothesis, I also asked respondents about government responsibility for "people confronting short-term hardships."  As Table \@ref(tab:tab-ui-hard-wmt) displays, we find no evidence of any treatment effect for any of these policy areas among Walmart workers; pre-exposure to the ACNT makes no difference in this conclusion.  Across both the Home Depot and Walmart surveys, we see no evidence that "private welfare" in the form of EHFs has any negative effect on support for government-directed social support, whether for childcare, the unemployed, the elderly, or those experiencing sudden hardship.

```{r tab-ui-hard-wmt, eval=TRUE, echo = FALSE, tab.cap="Treatment effect on support for unionization (Walmart)"}

ui_hard_tab_wmt
```


Findings related to the "substitution hypothesis"--the claim that private benefits and resources reduce support for government-provided programs--were another area of agreement across Studies 2 and 3.  Specifically, the video treatment shows no evidence of a negative effect on respondent support for government unemployment insurance or aid to those in temporary hardship.  In the Home Depot sample, the treatment induced a modest but significant _increase_ in support for government-provided UI.  This outcome may be the result of Home Depot workers having a better understanding of the EHF program (and its limitations).  Findings in these surveys are inconsistent with a simple "private options crowd out public benefits" story.



# References {-}

<div id="refs"></div>

\newpage

# (APPENDIX) Appendix {-}

\newpage

# Placebos for UI support (Home Depot) {#app-placebo}
```{r tab-placebo, echo=FALSE, tab.cap="Placebo regressions for UI"}

placebo_tab

```


